---
title: "2021_5_25"
author: "David Webber"
output: pdf_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
records <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-05-25/records.csv')
drivers <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-05-25/drivers.csv')
```

```{r}
records %>%
  group_by(player) %>%
  summarize(count = n())
```

```{r}
records %>%
  filter(track == "Rainbow Road", type == "Single Lap", system_played == "PAL", shortcut == "Yes")
  #ggplot(aes(x = date, y = time)) +
    #geom_point()
```

How many tracks have shortcuts? 12 tracks have shortcuts on them. The ones that don't have shortcuts are Banshee Boardwalk, Bowser's Castle, Koopa Troopa Beach, Moo Moo Farm.

```{r}
shortcuts <- records %>%
  filter(shortcut == "Yes") %>%
  group_by(track) %>%
  summarize(count = n())

shortcut_tracks <- unique(shortcuts$track)

records %>%
  filter(!(track %in% shortcut_tracks)) %>%
  group_by(track) %>%
  summarize(count = n())

```

Which track is the fastest? Mario Raceway is the fastest track for speedrunners on average while Rainbow Road has the longest average completion time.

```{r}
records %>%
  group_by(track) %>%
  summarize(ave_time = mean(time)) %>%
  arrange(ave_time)
```

Fastest times ever? Wario stadium has the fastest times of any track.
```{r}
records %>%
  arrange(time)
```

```{r}
#Function for plots

wr_plot <- function(a,b,c,d) {
  records %>%
    filter(track == a, type == b, system_played == c, shortcut == d) %>%
    ggplot(aes(date, time)) +
      geom_point()
}

wr_plot("Luigi Raceway", "Single Lap", "PAL", "Yes")
```

